---
title: "Linear Models: Homework 2"
author: "Your Name"
date: "2024-2025"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Answers to the questions

## Question 1a

Concise, to-the-point answer to question 1a

## Question 1b

Concise, to-the-point answer to question 1b

## Question 1c

Concise, to-the-point answer to question 1c

## Question 1d

Concise, to-the-point answer to question 1d

## Question 1e

Concise, to-the-point answer to question 1e

## Question 1f

Concise, to-the-point answer to question 1f

## Question 1g

Concise, to-the-point answer to question 1g


\newpage

# Appendix with R code

```{r}
# Change the path to the data file in the following line
load(file="~/dropbox/education/LIMO/homeworks/2425/HW2/exam.RData")

str(exam)

exam$Biostat<-as.numeric(exam$specialisation=="BS")
exam$DataSc<-as.numeric(exam$specialisation=="D")
exam$Bioinf<-as.numeric(exam$specialisation=="BI")
exam$Epi<-as.numeric(exam$specialisation=="E")
```

```{r}
m1<-lm(exam~homeworks+Biostat+Epi+Bioinf,data=exam)
summary(m1)
```

