---
title: "Linear Models"
author: "Your Name"
date: "20/10/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

This is a template R Markdown file. 

We advise you to repeat the examples and make the exercises in your own local R Markdown or Notebook file. You may use this file as a template, but use your own expertise and imagination to make it happen. 
You will never be asked to share this file with the lecturers. 

In the next section I've copied-and-pasted some of the R code from the html course notes on GitHub. 

# Galton's example

Make sure that you download the data and that the paths (directories) to the files are set correctly at your local computer. Adopt the habbit to comment the R code yourself and formulate conclusions, refer to course notes .... whatever you think is meaningful and helpful for you. 

The next few lines will read the Galton data from the source data file. 
```{r}
Galton<-read.csv("Data/Galton.tab",sep="\t")
head(Galton)
```

Next we select the relevant data and ...

```{r}
Galton.sons<-Galton %>%
  filter(gender=="M") %>%
  group_by(family) %>%
  slice(1) %>%
  ungroup %>%
  mutate(father.cm=father*2.54,
         son.cm=height*2.54) %>%
  select(father.cm, son.cm)
glimpse(Galton.sons)
```

Let's graphically explore the data ...

```{r}
ggplot(Galton.sons,
       aes(x=father.cm, y=son.cm)) +
  geom_point(color="blue") +
  xlab("length of father (cm)") + ylab("length of son (cm)") +
  theme(axis.title=element_text(size=15), axis.text = element_text(size=15))
```



